A Java Wrapper For Embarrassingly Parallel Problems

### 2013-11-02 paper/paper.tex
Graphs and table and more revisions

A computation graph consists of nodes that represent the functions or processes 
and the data being passed from one process to another is displayed in the graph 
as edges. If the graph is disconnected, none of the processes need to 
communicate to complete the computations. Such a computation is called
"embarrassingly parallel." In a practical situation, there will be some 
internode communication to set up the problem and then again to accumulate the 
results back together. Also known as pleasingly parallel problems, they can 
easily be performed on a cluster of servers or in any kind of distributed 
network, due to its lack of dependencies. The term  embarrassingly parallel was 
coined by FINISH THIS

@@

The framework provides a "wrapper' for the original program that mimics the
behavior of the original program. It calls the serial programs on multiple Java 
threads. Each thread acts like the serial application, reading a subset of the 
input the same and creating the output in the same format. This paper describes
a framework a framework that makes it easy to parallelize an embarrassingly
parallel program that reads from a file, writes to a file and whose input can be
broken up into work units by a regular expression. SUMMARY OF RESULTS to end the
paragraph.

@@

REDO THIS PART
The Ka/Ks Calculator is not the only embarrassingly parallel program that has the
potential to take several days to run on a data set. In fact, there are several
embarrassingly parallel programs in the biological sciences. The Basic Local
Alignment Search Tool (BLAST) \cite{BLAST} is a common program used to find
similarities between sequences and then compares them to databases to calculate
the statistical significance. Another program we have been looking at is called
Trinity. Trinity claims to be an efficient way to do de novo reconstruction of
transcriptomes from mRNA sequence data. With this variety of applications that 
would benefit from parallelism, it was obvious it would be desirable to develop
a framework for creating wrapper programs.

@@

Table and chart

The Ka/Ks Calculator was run through the framework with ten different samples,
each containing sixteen units of work. Each sample was run by the framework at 
five different level of numbers of worker threads, starting at 1 and 
increasing by powers of two until the 100\% CPU usage (16 processors). As the 
numbers of threads increased, the time it took descreased exponentially.

@@

In a few cases, there was an increase in time when attempting to max out the CPU 
usage. This occured because the framework had to wait until for the CPU to be 
finished with its previous work before it was allowed to start work on what the 
framework wanted it to do. In the cases when the framework executed faster while
the CPU was maxed out, it was only faster a minute at the most. With such a
small time gain with the CPU at 50\% versus the CPU at 100\%, running the
framework at 50\% CPU is more efficient when the number of threads is on the
same order as the number of units of work.

@@

Using Java was a good choice for implementation ease and future implementation
extension by others, but it may not produce the fastest results. Languages
like Hadoop, Clojure or other multithread oriented language may have been better
due to their parallel nature. Another approach that could have been taken is one
similar to the people in Washington: going to the cloud. Using a networked 
solution like MPI or a cloud solution like Amazon instances. (More will go here
when I added in more previous work from others)

\begin{thebibliography}{1}
\bibitem{kaks}
Zhang Z, Li J, Zhao XQ, Wang J, Wong GK, Yu J., ``\emph{KaKs Calculator: 
calculating Ka and Ks through model selection and model averaging.},'',
Genomics Proteomics Bioinformatics, November 2006.
\bibitem{BLAST}
http://blast.ncbi.nlm.nih.gov/Blast.cgii
\end{thebibliography}

>>>>

A computation graph consists of nodes that represent the functions or processes 
and the data being passed from one process to another is displayed in the graph 
as edges. If the graph is disconnected, none of the processes need to 
communicate to complete the computations. Such a computation is called
"embarrassingly parallel." In a practical situation, there will be some 
internode communication to set up the problem and then again to accumulate the 
results back together. Also known as pleasingly parallel problems, they can 
easily be performed on a cluster of servers or in any kind of distributed 
network, due to its lack of dependencies. The term embarrassingly parallel was 
coined by Cleve Moler in his 1986 paper "Matrix Computation on Distributed
Memory Multiprocessors." \cite{history}

@@

The framework provides a "wrapper' for the original program that mimics the
behavior of the original program. It calls the serial programs on multiple Java 
threads. Each thread acts like the serial application, reading a subset of the 
input the same and creating the output in the same format. This paper describes
a framework a framework that makes it easy to parallelize an embarrassingly
parallel program that reads from a file, writes to a file and whose input can be
broken up into work units by a regular expression. The framework gives an
exponential decrease in runtime as more processor help in the computation. This
decrease levels off between 50 and 100\% CPU usage due to the other tasks on the
machine that take up memory.

@@

REDO THIS PART
The Ka/Ks Calculator is not the only embarrassingly parallel program that has the
potential to take several days to run on a data set. In fact, there are several
embarrassingly parallel programs in the biological sciences. 
TALK ABOUT CLUSTERING
With several applications that 
would benefit from parallelism in this field, it was obvious it would be 
desirable to develop a framework for creating wrapper programs.

@@

To observe how the framework changes the runtime of sequential programs when
using varying numbers of threads, I designed a simple experiment. The Ka/Ks 
Calculator was run through the framework with ten different samples, each 
containing sixteen units of work. Each sample was run by the framework at five 
different level of numbers of worker threads, starting at 1 and increasing by 
powers of two until the 100\% CPU usage (16 processors). These tests were run
with minimal other activity on the server.

\begin{figure}
{{!}{\includegraphics{experiment/out.png}}}
\caption{Runtimes for the Ka/Ks Calculator though the framework}
\label{fig:graph}
\end{figure}

@@

As the numbers of threads increased, the time it took descreased exponentially. 
In a few cases, there was an increase in time when attempting to max out the CPU 
usage. This occured because the framework had to wait until for the CPU to be 
finished with its previous work before it was allowed to start work on what the 
framework wanted it to do. In the cases when the framework executed faster while
the CPU was maxed out, it was only faster a minute at the most. With such a
small time gain with the CPU at 50\% versus the CPU at 100\%, running the
framework at 50\% CPU is more efficient when the number of threads is on the
same order as the number of units of work.

@@

Using Java was a good choice for implementation ease and future implementation
extension by others, but it may not produce the fastest results. Languages
like Hadoop, Clojure or other multithread oriented language may have been better
due to their parallel nature. In addition to a single machine working to process
the data quickly, we can also look to the cloud for help by using a networked 
solution like MPI or a cloud solution like Amazon instances. This would
potentially allow for an even greater time decrease because the number of
machines that could help is most likely larger than the number of processors on
one machine.

\begin{thebibliography}{1}
\bibitem{history}
Moler, Cleve. "Matrix Computation on Distributed Memory Multiprocessors."
Hypercube Multiprocessors, 1986: Proceedings of the First Conference on
Hypercube Multiprocessors, Knoxville, Tennessee, August 24-27, 1985. By Michael
T. Heath. Philadelphia: SIAM, 1986. N. pag. Print.
\bibitem{kaks}
Zhang Z, Li J, Zhao XQ, Wang J, Wong GK, Yu J., ``\emph{KaKs Calculator: 
calculating Ka and Ks through model selection and model averaging.},'',
Genomics Proteomics Bioinformatics, November 2006.
\end{thebibliography}
